---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# net4routing

<!-- badges: start -->
<!-- badges: end -->

The goal of net4routing is to prepare route networks, imported from OpenStreetMap or other sources, for routing with `cppRouting` and other packages.

## Installation

You can install the development version of net4routing from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("r-transport/net4routing")
```

## Example

```{r devtools, include = FALSE}
devtools::load_all()
```

```{r example}
library(net4routing)
# Download a network:
region_name = "isle of wight"
pbf_info = osmextract::oe_match(region_name)
pbf_url = pbf_info$url
pbf_file = osmextract::oe_download(pbf_info$url)
res = nr_osm4routing(pbf_file)
```

We'll now make a graph with cppRouting:

```{r graph}
library(cppRouting)
# Load the network
nodes = readr::read_csv(res[1]) |>
  dplyr::transmute(
    ID = id, X = lon, Y = lat
  ) |>
  as.data.frame()
edges = readr::read_csv(res[2]) |>
  dplyr::transmute(
    from = source, to = target, weight = length
  ) |>
  as.data.frame()
graph = cppRouting::makegraph(edges, coords = nodes)
str(graph)
# List of 5
#  $ data  :'data.frame': 51940 obs. of  3 variables:
#   ..$ from: int [1:51940] 0 1 0 2 3 4 5 6 7 8 ...
#   ..$ to  : int [1:51940] 1 211 2 3 4 5 6 7 8 9 ...
#   ..$ dist: num [1:51940] 14.6 39.1 104.1 43.5 41.5 ...
#  $ coords:'data.frame': 42852 obs. of  3 variables:
#   ..$ ID: chr [1:42852] "255734" "5339263635" "255733" "765635519" ...
#   ..$ X : num [1:42852] -1.08 -1.08 -1.08 -1.08 -1.08 ...
#   ..$ Y : num [1:42852] 50.7 50.7 50.7 50.7 50.7 ...
#  $ nbnode: int 42852
#  $ dict  :'data.frame': 42852 obs. of  2 variables:
#   ..$ ref: chr [1:42852] "255734" "5339263635" "255733" "765635519" ...
#   ..$ id : int [1:42852] 0 1 2 3 4 5 6 7 8 9 ...
#  $ attrib:List of 4
#   ..$ aux  : NULL
#   ..$ cap  : NULL
#   ..$ alpha: NULL
#   ..$ beta : NULL
```

You can now use this graph to find the shortest path between two points:

```{r shortest_path}
# Find the shortest path between two points
from = "255734"
to = "255738"
shortest_path = cppRouting::get_path_pair(
  graph,
  from = from,
  to = to
)
shortest_path
# $`255734_255738`
#  [1] "255734"      "255733"      "765635519"   "279387"      "540707"     
#  [6] "3792504678"  "10760063960" "10760063953" "4580466911"  "10760063950"
# [11] "2306536"     "2306537"     "2387732"     "4339515020"  "255738"  
# Check all are in the edge list
summary(shortest_path[[1]] %in% nodes$ID)
nodes_edges = nodes[nodes$ID %in% shortest_path[[1]], ]
# Plot them:
plot(
  nodes_edges$X, nodes_edges$Y,
  pch = 19, col = "blue", cex = 0.5,
  xlab = "Longitude", ylab = "Latitude"
)
```

```{r oan}
set.seed(123)
demand = data.frame(
  from = sample(nodes$ID, 5000, replace = TRUE),
  to = sample(nodes$ID, 5000, replace = TRUE),
  demand = runif(5000, 1, 10)
)
aon = cppRouting::get_aon(
  graph,
  from = demand$from,
  to = demand$to,
  demand = demand$demand,
)
head(aon)
```
# Development

To format code for the package, install and format with `air`:

```sh
curl -LsSf https://github.com/posit-dev/air/releases/latest/download/air-installer.sh | sh
air format .
```

Check the package with:

```{r check, eval=FALSE}
devtools::check()
```